---
title: "p8105_mtp_dtw2127"
author: "Dee Wang"
date: "27/10/2021"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Introduction 

We examine the findings presented in the 2018 article on the relationship between posture and enlarged external occipital protuberances. We examine data consistency, results presentation and try and reproduce results. 

## Problem 1 - Data
We fill in missing values for eop_size_mm with 0s, convert categorical variables into character and ordered factors, and recode age_group and sex according to header information.

```{r}
nature_data = read_excel("./data/p8105_mtp_data.xlsx", sheet = "this one", skip = 8) %>%
  janitor::clean_names() %>% 
  mutate(sex = as.factor(sex),
         age_group = as.factor(age_group),
         eop_size = as.factor(eop_size), 
         eop_shape = as.factor(eop_shape),
         fhp_category = as.factor(fhp_category))

#replace missing values in eop_size_mm column with 0s'
nature_data[["eop_size_mm"]][is.na(nature_data[["eop_size_mm"]])] <- 0 

nature_data = nature_data %>%
  filter(as.numeric(age_group) > 1) %>% #filter out age_group 1
  mutate(age_group = recode(age_group,
                            `2` = "18-30 yrs", 
                            `3` = "31-40 yrs",
                            `4` = "41-50 yrs",
                            `5` = "51-60 yrs",
                            `6` = "60+ yrs",
                            `7` = "60+ yrs",
                            `8` = "60+ yrs")) %>%
  mutate(sex = recode(sex, 
                      `0` = "female",
                      `1` = "male")) 

```

Key variables in the dataset include sex and age, eop_size, eop_visibility_classification, eop_shape and fhp_size. There are `r nrow(nature_data)` participants included in the dataset. 

```{r}
  nature_data %>% 
    group_by(age_group, sex) %>%
  summarize(n_obs = n()) %>%
pivot_wider(
  names_from = age_group,
  values_from = n_obs)

```

The 18-30, and 60+ years old are the largest groups. There are roughly equal numbers of males and females. 

We'll check categorical variables assignment.

```{r}

nature_data_check = nature_data %>% 
  mutate(correct_age_category = 
           case_when(age >= 18 & age <= 30 ~ "18-30 yrs", 
                     age >= 31 & age <= 40 ~ "31-40 yrs", 
                     age >= 41 & age <= 50 ~ "41-50 yrs", 
                     age >= 51 & age <= 60 ~ "51-60 yrs", 
                     age >= 61 ~ "60+ yrs"), 
         age_check = ifelse(correct_age_category == age_group, "consistent", "inconsistent"), 
         correct_size_category = 
           case_when(eop_size_mm < 5 ~ 0, 
                     eop_size_mm >= 5 & eop_size_mm < 10 ~ 1, 
                     eop_size_mm >= 10 & eop_size_mm < 15 ~ 2, 
                     eop_size_mm >= 15 & eop_size_mm < 20 ~ 3,
                     eop_size_mm >= 20 & eop_size_mm <= 25 ~ 4,
                     eop_size_mm >= 25 ~ 5),
         size_check = ifelse(correct_size_category == eop_size, "consistent", "inconsistent"),
         correct_vis_class = 
           case_when(eop_size_mm == 0 ~ 0, 
                     eop_size_mm > 0 & eop_size_mm <= 5 ~ 1, 
                     eop_size_mm > 5 ~ 2), 
         vis_check = ifelse(correct_vis_class == eop_visibility_classification, "consistent", "inconsistent"),
         correct_fhp_category = 
           case_when(fhp_size_mm < 10 ~ 0,
                     fhp_size_mm >= 10 & fhp_size_mm < 20 ~ 1, 
                     fhp_size_mm >= 20 & fhp_size_mm < 30 ~ 2, 
                     fhp_size_mm >= 30 & fhp_size_mm < 40 ~ 3,
                     fhp_size_mm >= 40 & fhp_size_mm < 50 ~ 4, 
                     fhp_size_mm >= 50 & fhp_size_mm < 60 ~ 5, 
                     fhp_size_mm >= 60 & fhp_size_mm < 70 ~ 6, 
                     fhp_size_mm >= 70 & fhp_size_mm < 80 ~ 7, 
                     fhp_size_mm >= 80 & fhp_size_mm < 90 ~ 8), 
         fhp_check = ifelse(correct_fhp_category == fhp_category, "consistent", "inconsistent")
         )
 

```


```{r}

nature_data_check %>% 
  select(age, age_group, correct_age_category, age_check) %>%
  filter(age_check == "inconsistent")

```

Age groups were correctly assigned. 


```{r}
nature_data_check %>% 
  select(eop_size_mm, eop_size, correct_size_category, size_check) %>% 
  filter(size_check == "inconsistent")
```

`r nature_data_check %>% select(eop_size_mm, eop_size, correct_size_category, size_check) %>% filter(size_check == "incorrect") %>% nrow()` participants had incorrect EOP size categorizations. Many participants were assigned EOP size 4 instead of 5. 

```{r}
nature_data_check %>% 
  select(eop_size_mm, eop_visibility_classification, correct_vis_class, vis_check) %>% 
  filter(vis_check == "inconsistent") 
```

`r nature_data_check %>% 
  select(eop_size_mm, eop_visibility_classification, correct_vis_class, vis_check) %>% 
  filter(vis_check == "incorrect") %>% nrow()` participants had their visibility classification incorrectly assigned. Many participants were categorized as EOP size 0 instead of 1.

```{r}
nature_data_check %>% 
  select(fhp_size_mm, fhp_category, correct_fhp_category, fhp_check) %>%
  filter(fhp_check == "inconsistent") 
```
`r nature_data_check %>% select(fhp_size_mm, fhp_category, correct_fhp_category, fhp_check) %>% filter(fhp_check == "inconsistent") %>% nrow()` individuals had an incorrect FHP categorization (FHP category underestimated). 

Number of data inconsistencies is minimal, so we'll leave this data in. 

## Problem 2 - Visualization 

```{r}

figure_3 = ggplot(nature_data,
                  aes(x = age_group, y = fhp_size_mm, fill = sex)) +
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "age group", y = "FHP size (mm)")

```

```{r}
#figure out how many participants are in each age + sex group
nature_data_age_n = nature_data %>% 
  group_by(age_group, sex) %>% 
  summarize(n_sex_age = n())

#figure out how many participants within each age+sex group have EEOP.
nature_data_eop_n = nature_data %>% 
  filter(eop_size_mm >= 10) %>%
  group_by(age_group, sex) %>% 
  summarize(n_eeop = n()) 

nature_data_combined = 
  left_join(nature_data_age_n, nature_data_eop_n) %>% 
  mutate(rate = n_eeop/n_sex_age) %>% 
  mutate(rate = round(rate, 2))

figure_4 = 
  nature_data_combined %>% 
  ggplot(aes(x = age_group, y = rate, color = sex)) + 
  geom_line(aes(group = sex)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "age group", y = "EEOP rate")

(figure_3 + figure_4)

```

From the right panel, We can see a higher rate of EEOP in males. The highest rates of EEOP are in the youngest age group. If FHP is predictive of EEOP, we would expect the FHP size trend on the left to be similar to that of the EEOP rate trend, but this is not the case. While FHP sizes appear to be larger in males as compared to females, it isn't clear if the differences are significant. 

```{r}

nature_data %>% 
  #filter(eop_size_mm > 0) %>% 
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm)) +
  geom_point(size = .3) + 
  labs() + 
  facet_grid(sex ~ age_group)
  
```

There is no association between EOP size and FHP size in any of the age and sex groups. Plots don't provide evidence that FHP is predictive of EOP.

## Problem 3 - Reproducing reported results 

We'll try reproducing some reported results.

```{r}
nature_data %>% 
  count(age_group, name = "n_age_group")
```

Sample sizes calculated are similar to those reported by the authors (differences within 7). 

```{r}
nature_data %>% 
  group_by(sex) %>% 
  summarize(mean_fhp = mean(fhp_size_mm, na.rm = TRUE), 
            sd_fhp = sd(fhp_size_mm, na.rm = TRUE)) %>%
  mutate(mean_fhp = round(mean_fhp, digits = 2), 
         sd_fhp = round(sd_fhp, digits = 2))

```

Authors' calculation of mean and standard deviations for FHP size are consistent with our calculations. 

The definition of EEOP is an EOP size > 10 mm. We'll use the EOP size variable to check the EEOP prevalence calculation.

```{r}

eeop_n = nature_data %>%
  filter(as.numeric(eop_size_mm) >= 10) %>% 
  nrow()

study_pop_n = nrow(nature_data)

```
The prevalence we calculate is `r eeop_n/study_pop_n` is similar to the authors' calculations, but ~1% off. 

```{r}
nature_data_age_n = nature_data %>%
  group_by(age_group) %>% 
  summarize(n_age_group = n())

nature_data_fhp_over40_n = nature_data %>%
  filter(fhp_size_mm > 40) %>%
  group_by(age_group) %>% 
  summarize(n_over40 = n())
  
table = left_join(nature_data_age_n, nature_data_fhp_over40_n) %>% 
  mutate(prevalence = n_over40/n_age_group*100) %>% 
  mutate(prevalence = round(prevalence, 2))

print(table)

```
Prevalence of FHP over 40 mm is higher in older age groups and highest in the 60+ years group which is consistent with authors' observations. Our prevalence calculation for 60+ years is different by ~2%. 

## Problem 4 - Discussion 

There are issues with the data analyses and result presentation. Some categorizations did not match data header information. Figures were also misleading, especially Figure 4. We calculated similar results as reported in the paper, but couldn't get the exact calculations. 

Based on this paper, we cannot conclude that cell phones are causing horn growth. We would need to see data on cell phone usage, and evidence that screen-based activity impacts FHP size. 

### Word count
```{r}
wordcountaddin::text_stats("p8105_mtp_dtw2127.Rmd")
```




